On exponential sums arising from the classical doubling method

### 2025-02-07 paper.tex
Started writing about Saito-Shinoda

\newcommand{\parabolicSection}{\Phi^{\left(z_1, \dots, z_c\right)}}
\newcommand{\intertwiningOperator}{M^{\left(z_1, \dots, z_c\right)}}
\newcommand{\holomorphicRepresentation}{\depthZeroRepresentation^{\left(z_1, \dots, z_c\right)}}
\newcommand{\WhittakerFunctional}[1]{\ell_{#1, \fieldCharacter}}
\newcommand{\gWhittakerFunctional}[2]{\ell_{#1, #2}}
\newcommand{\gSpehWhittakerFunctional}[3]{\ell_{\SpehRepresentation{#1}{#3}, \fieldCharacter_{\UnipotentRadicalForWss{#2}{#3}}}}
\newcommand{\gShortSpehWhittakerFunctional}[3]{\ell_{\SpehRepresentation{#1}{#3}}}
\newcommand{\GaussSum}[2]{\mathcal{G}\left(#1, #2\right)}
\newcommand{\dblGaussSum}[2]{\mathcal{G}^{\mathrm{dbl}}\left(#1, #2\right)}
\newcommand{\GaussSumScalar}[2]{\mathrm{G}\left(#1, #2\right)}
\newcommand{\dblGaussSumScalar}[2]{\mathrm{G}^{\mathrm{dbl}}\left(#1, #2\right)}
\newcommand{\dblGammaFactor}[3]{\Gamma^{\mathrm{dbl}}\left(#1 \times #2, #3\right)}
\newcommand{\dblGammaFactorSpace}[4]{\Gamma^{\mathrm{dbl}}_{#1}\left(#2 \times #3, #4\right)}
\newcommand{\GKGaussSum}[3]{\mathcal{G}\left(#1 \times #2, #3\right)}
\newcommand{\GKGaussSumScalar}[3]{\mathrm{G}\left(#1 \times #2, #3\right)}
\newcommand{\fieldCharacterkc}[2]{\fieldCharacter_{\left({#2}^{#1}\right)}}
\newcommand{\ExoticKloosterman}{\mathrm{Kl}}
\newcommand{\GaussSumCharacter}[4]{\tau_{#1}\left(#2 \times #3, #4\right)}
\newcommand{\IrrCusp}{\Irr_{\mathrm{cusp}}}
\newcommand{\convolutionWithCompactSupport}{\boldsymbol{\mathrm{R}}}
\newcommand{\ladicnumbers}{\algebraicClosure{\mathbb{Q}_{\ell}}}
\newcommand{\artinScrier}{\operatorname{AS}}
\newcommand{\KloostermanSumClassFunction}{\mathcal{K}}
\newcommand{\IsometryGroup}{\mathrm{Isom}}
\newcommand{\lieAlgebra}{\mathfrak{g}}
\newcommand{\DeligneLusztigInduction}[2]{\mathrm{R}_{#1}^{#2}}

@@

>>>>

\newcommand{\parabolicSection}{\Phi^{\left(z_1, \dots, z_c\right)}}
\newcommand{\intertwiningOperator}{M^{\left(z_1, \dots, z_c\right)}}
\newcommand{\holomorphicRepresentation}{\depthZeroRepresentation^{\left(z_1, \dots, z_c\right)}}
\newcommand{\WhittakerFunctional}[1]{\ell_{#1, \fieldCharacter}}
\newcommand{\gWhittakerFunctional}[2]{\ell_{#1, #2}}
\newcommand{\gSpehWhittakerFunctional}[3]{\ell_{\SpehRepresentation{#1}{#3}, \fieldCharacter_{\UnipotentRadicalForWss{#2}{#3}}}}
\newcommand{\gShortSpehWhittakerFunctional}[3]{\ell_{\SpehRepresentation{#1}{#3}}}
\newcommand{\GaussSum}[2]{\mathcal{G}\left(#1, #2\right)}
\newcommand{\dblGaussSum}[2]{\mathcal{G}^{\mathrm{dbl}}\left(#1, #2\right)}
\newcommand{\GaussSumScalar}[2]{\mathrm{G}\left(#1, #2\right)}
\newcommand{\dblGaussSumScalar}[2]{\mathrm{G}^{\mathrm{dbl}}\left(#1, #2\right)}
\newcommand{\dblVirtualGaussSumScalar}[2]{\mathrm{g}^{\mathrm{dbl}}\left(#1, #2\right)}
\newcommand{\dblGammaFactor}[3]{\Gamma^{\mathrm{dbl}}\left(#1 \times #2, #3\right)}
\newcommand{\dblGammaFactorSpace}[4]{\Gamma^{\mathrm{dbl}}_{#1}\left(#2 \times #3, #4\right)}
\newcommand{\GKGaussSum}[3]{\mathcal{G}\left(#1 \times #2, #3\right)}
\newcommand{\GKGaussSumScalar}[3]{\mathrm{G}\left(#1 \times #2, #3\right)}
\newcommand{\fieldCharacterkc}[2]{\fieldCharacter_{\left({#2}^{#1}\right)}}
\newcommand{\ExoticKloosterman}{\mathrm{Kl}}
\newcommand{\GaussSumCharacter}[4]{\tau_{#1}\left(#2 \times #3, #4\right)}
\newcommand{\IrrCusp}{\Irr_{\mathrm{cusp}}}
\newcommand{\convolutionWithCompactSupport}{\boldsymbol{\mathrm{R}}}
\newcommand{\ladicnumbers}{\algebraicClosure{\mathbb{Q}_{\ell}}}
\newcommand{\artinScrier}{\operatorname{AS}}
\newcommand{\KloostermanSumClassFunction}{\mathcal{K}}
\newcommand{\IsometryGroup}{\mathrm{Isom}}
\newcommand{\lieAlgebra}{\mathfrak{g}}
\newcommand{\DeligneLusztigInduction}[2]{\mathrm{R}_{#1}^{#2}}
\newcommand{\algebraicGroup}[1]{\boldsymbol{\mathrm{#1}}}

@@

\subsection{Computation in terms of Deligne--Lusztig data}

The goal of this section is to compute $\dblGaussSumScalar{\pi}{\chi}$ for $\pi$ that appears in a Deligne--Lusztig virtual character $R_{T,\theta}$ for a suitable maximal torus $T$ and $\theta \colon T \to \multiplicativegroup{\cComplex}$.

Let us realize $G$ as the fixed points of the Frobenius map $\Frobenius$ acting on a connected reductive algebraic group $\algebraicGroup{G}$.

Given a class function $F \colon G \to \cComplex$ (i.e., a function that is invariant under conjugation by elements of $G$), let us define $$\dblVirtualGaussSumScalar{F}{\chi} = \frac{1}{\sqrt{\sizeof{\lieAlgebra}}} \sum_{g \in G} F\left(g\right) \Phi_{\chi}\left(g\right).$$

It is clear that if $\pi$ is an irreducible representation of $G$ then $$\dblVirtualGaussSumScalar{\trace \pi}{\chi} = \dblGaussSumScalar{\pi}{\chi} \cdot \dim \pi.$$
Notice that the assignment $g \mapsto \Phi_{\chi}\left(g\right)$ only depends on the semisimple part of $g \in G$. By \cite[Theorem in Section 1.2]{SaitoShinoda2000}, we have the following result.
\begin{proposition}
	For any $\Frobenius$-stable maximal torus $\algebraicGroup{T}$ of $\algebraicGroup{G}$, and any character $\theta \colon T \to \multiplicativegroup{\cComplex}$, where $T = \algebraicGroup{T}^{\Frobenius}$, we have
	$$ \dblVirtualGaussSumScalar{R_{T, \theta}}{\chi} = \frac{\grpIndex{G}{T}}{\sqrt{\sizeof{\lieAlgebra}}} \sum_{t \in T} \theta\left(t\right) \Phi_{\chi}\left(t\right).$$
\end{proposition}
In particular, if $\theta$ is in general position, then $\varepsilon_{\algebraicGroup{G}} \varepsilon_{\algebraicGroup{T}} R_{T, \theta}$ equals $\trace \pi$ for some irreducible representation $\pi$ of $G$ and we have that for this $\pi$,
$$\dblGaussSumScalar{\pi}{\chi} = \varepsilon_{\algebraicGroup{G}} \varepsilon_{\algebraicGroup{T}} \sizeof{G}_p \sum_{t \in T} \theta\left(t\right) \Phi_{\chi}\left(t\right).$$
Here $\varepsilon_{\algebraicGroup{G}} = \left(-1\right)^{\mathrm{rel.rank} \algebraicGroup{G}}$ and $\varepsilon_{\algebraicGroup{T}} = \left(-1\right)^{\mathrm{rel.rank} \algebraicGroup{T}}$, and $\sizeof{G}_p$ is the size of the $p$-Sylow group of $G$, where $p$ is the characteristic of $\finiteField$.
